Future Cosmic Microwave Background delensing with galaxies surveys.

### 2017-06-03 Notes/DES_rec_phi_paper.tex
work on writing

Delensing or more, in general, the ability to separate the lensing component of the B-mode from a possible primordial inflationary signal is important to exploit the capabilities of future experiments fully.
In this paper, we studied the potential impact of large-scale structure surveys in this important endeavor.
For current experiment, CIB data had already proven to be very efficient in delensing. If no high-resolution CMB experiment is available to reconstruct the lensing potential $\phi$ a low-z optical experiment like D.E.S will improve the delensign efficiency by $~10\%$ if combined with Planck lensing and a CIB tracers.
As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential. Indeed using the CMB itself, we reconstruct only the lenses that lens the CMB obtained an almost perfect cleaning in the absence of noise. This approach will need a careful study of possible biases coming from using the same source (the CMB) we are delensing to reconstruct the lensing effect itself. This has already been applied to data and studied. However, the relatively good efficiency of galaxies tracers might be useful to cross-check these internal biases. This will probably be needed to confirm a possible detection of gravitation waves if this relies heavily on delensing.

Another possible application of LSS delensing is the removal of the low-z lensing of the CMB that possibly adds the complexity of modeling nonlinearities in the study of CMB lensing

>>>>

Delensing or the ability to separate the lensing component of the CMB B-mode from a possible primordial inflationary signal is important to exploit the capabilities of future experiments thoroughly.
In this paper, we studied the potential impact of large-scale structure surveys in this important endeavor.
For current experiment, CIB data had already proven to be very efficient in delensing \cite{larsen:2016,manzotti:2017}. 
If no CMB experiment is available with higher resolution then Planck is used to reconstruct the lensing potential $\phi$, a low-z optical survey like D.E.S will improve the delensing efficiency by $~10\%$ when combined with Planck lensing and a CIB tracers.
As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential. Indeed using the CMB itself, we reconstruct only the lenses that lens the CMB obtaining, as a result, an almost perfect cleaning in the absence of noise. 
This approach has already been applied to data \cite{carron:2017}. However it requires a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}. For this reason, the relatively good efficiency of galaxies tracers might be useful to cross-check these internal biases. This will probably be needed to confirm a possible detection of gravitation waves if this relies heavily on delensing.

Another possible application of LSS delensing, which we will explore in future works, is the possibility of removing just the low-z lensing of the CMB which will otherwise possibly adds the complexity of modeling nonlinearities in the study of CMB lensing.
